---
title: " Replication of College Football Games and Crime (Rees & Schnepel, 2009)"
author: "Jacqueline and Farida"
date: "`r Sys.Date()`"
output: html_document
---

```{r load-packages}
library(cfbfastR)
library(tidyverse)
library(readr)
library(here)
library(purrr)
library(httpgd)

theme_set(theme_bw())
here()

knitr::opts_chunk$set(echo = TRUE)
```

## Loading all the data

```{r}
source(here("scripts/03_load_offense_data.R"))
source(here("scripts/04_load_football_college.R"))
```

## Replicating core analysis

This is a template that aims to replicate and analyze Daniel Rees and Kevin Schnepel's paper, College Football Games and Crime published in 2009 by Journal of Sports Economics. 
This paper (https://jvlone.com/sportsdocs/footballGamesCrime2009.pdf) examines whether college football games lead to an increase in crime in host communities. 

### Descriptive statistics: Reproduce Appendix Table 1 from the paper: compute the mean, standard deviation, and quartiles of daily assault and vandalism counts across your sample of college town agencies. 

#### All Games Descriptive Statistics Count
```{r}
final_table %>%
    summarize(
        mean = mean(assault), 
        sd = sd(assault),
        quartile_25 = quantile(assault, 0.25),
        quartile_50 = quantile(assault, 0.5),
        quartile_75 = quantile(assault, 0.75),
        quartile_90 = quantile(assault, .90)
    )

final_table %>%
    summarize(
        mean = mean(vandalism), 
        sd = sd(vandalism),
        quartile_25 = quantile(vandalism, 0.25),
        quartile_50 = quantile(vandalism, 0.5),
        quartile_75 = quantile(vandalism, 0.75),
        quartile_90 = quantile(vandalism, .90)
    )
```

#### No Games Descriptive Statistics Count

```{r}
no_game_table <- filter(final_table, home_or_away == "none")
no_game_table %>%
    summarize(
        mean = mean(assault), 
        sd = sd(assault),
        quartile_25 = quantile(assault, 0.25),
        quartile_50 = quantile(assault, 0.5),
        quartile_75 = quantile(assault, 0.75),
        quartile_90 = quantile(assault, .90)
    )

no_game_table %>%
    summarize(
        mean = mean(vandalism), 
        sd = sd(vandalism),
        quartile_25 = quantile(vandalism, 0.25),
        quartile_50 = quantile(vandalism, 0.5),
        quartile_75 = quantile(vandalism, 0.75),
        quartile_90 = quantile(vandalism, .90)
    )
```

#### Home Games Descriptive Statistics Count

```{r}
home_game_table <- filter(final_table, home_or_away == "home")
home_game_table %>%
    summarize(
        mean = mean(assault), 
        sd = sd(assault),
        quartile_25 = quantile(assault, 0.25),
        quartile_50 = quantile(assault, 0.5),
        quartile_75 = quantile(assault, 0.75),
        quartile_90 = quantile(assault, .90)
    )

home_game_table %>%
    summarize(
        mean = mean(vandalism), 
        sd = sd(vandalism),
        quartile_25 = quantile(vandalism, 0.25),
        quartile_50 = quantile(vandalism, 0.5),
        quartile_75 = quantile(vandalism, 0.75),
        quartile_90 = quantile(vandalism, .90)
    )
```

#### Away Games Descriptive Statistics Count

```{r}
away_game_table <- filter(final_table, home_or_away == "away")
away_game_table %>%
    summarize(
        mean = mean(assault), 
        sd = sd(assault),
        quartile_25 = quantile(assault, 0.25),
        quartile_50 = quantile(assault, 0.5),
        quartile_75 = quantile(assault, 0.75),
        quartile_90 = quantile(assault, .90)
    )

away_game_table %>%
    summarize(
        mean = mean(vandalism), 
        sd = sd(vandalism),
        quartile_25 = quantile(vandalism, 0.25),
        quartile_50 = quantile(vandalism, 0.5),
        quartile_75 = quantile(vandalism, 0.75),
        quartile_90 = quantile(vandalism, .90)
    )
```

### Table 2 Reproduction

``` {r}
plot <- final_table %>% mutate(assault = ifelse(is.na(assault),0,assault), 
vandalism  = ifelse(is.na(vandalism),0,vandalism ), 
home_or_away =  ifelse(is.na(home_or_away),"None",home_or_away ))

plot
```
``` {r table-2}
#distribution of games by day of the week
# day of week, games , obersvations 
dist_of_games <- plot %>% group_by(weekday ) %>% summarize(games = sum(home_or_away == "home")+sum(home_or_away == "away"),observations  = n())
total <- tibble(weekday = "Total", games = sum(dist_of_games$games),observations = sum(dist_of_games$observations))
dist_of_games <- bind_rows(dist_of_games,total)
dist_of_games
```

## Descriptive figures
### Create figures showing the mean number of offenses on Saturdays, broken out by whether a home game, away game or no game was played. Start with assualts and vandalism.

``` {r descriptive-figures}
#mean of number of offenses by home game away game or no game 
mean <-plot %>% filter(weekday == "Saturday") %>%  group_by(home_or_away) %>% summarize(mean_assault = mean(assault), mean_vandalism =  mean(vandalism) , mean = mean(vandalism+assault), se = sd(vandalism+assault)/sqrt(n()))
ggplot(mean, aes (x = home_or_away, y = mean )) +
geom_errorbar(aes(ymin = mean - se,ymax = mean +se )) + geom_point()+
xlab('') +
ylab('Average Offenses')
mean
# mean number per teams
mean_per_team<- plot %>% left_join(all_game_and_oris) %>% filter(weekday == "Saturday") %>%  group_by(home_or_away,school_name) %>% summarize(mean_assault = mean(assault), mean_vandalism =  mean(vandalism) , mean = mean(vandalism+assault), se = sd(vandalism+assault)/sqrt(n()))
ggplot(mean_per_team, aes (x = home_or_away, y = mean )) +
geom_errorbar(aes(ymin = mean - se,ymax = mean +se )) + geom_point()+ facet_wrap(~ school_name,scales = "free_y")+
#geom_errorbar(aes(ymin = mean - se,ymax = mean +se )) + geom_point()+ facet_wrap(~ school_name)+
xlab('') +
ylab('Average Offenses')

```

### Per-team plot: Show the same comparison separately for each college town / team instead of aggregated across all teams.
Create figures showing the mean number of offenses on Saturdays (with standard errors), broken out by whether a home game, away game, or no game was played.

```{r}
ori_by_school <- tibble(
    ori = c("AR0720100", "AR0160100", "CO0210100", "ID0010100", "ID0290500", "IA0850100", "IA0520200", "KS0230100", "MI8121800", "MI3336400", "MI3949900", "MI3759900", "OH0770100", "OH0050100", "OHCOP0000", "SC0390200", "SC0400100", "TN0750100", "TX2270100", "TX0610200", "TX1520000", "UT0030100", "UT0250600", "VA0600100", "WV0060200", "WV0310100"),
    school_name = c("University of Arkansas", "Arkansas State University", "United States Air Force Academy", "
    Boise State University", "The University of Idaho", "Iowa State University", "The University of Iowa", "The University of Kansas", "The University of Michigan", "Michigan State University", "Western Michigan University", "Eastern Michigan University", "The University of Akron", "Ohio University", 
    "The Ohio State University", "Clemson University", "The University of South Carolina", "Middle Tennessee State University", "The University of Texas at Austin", "The University of North Texas", "Texas Tech University", "Utah State University", 
    "Brigham Young University", "Virginia Polytechnic Institute and State University", "Marshall University", "West Virginia University")
)

final_table_saturdays_only <- filter(final_table, weekday == "Saturday")

per_plot_team <- final_table_saturdays_only %>% 
    mutate(offenses = assault + vandalism) %>% 
    group_by(ori, home_or_away) %>% 
    summarize(mean = mean(offenses), se = sd(offenses) / sqrt(n()), .groups = "drop") %>%
    left_join(ori_by_school, by = "ori")

ggplot(per_plot_team, mapping = aes(x = home_or_away, y = mean, ymin = mean - se, ymax = mean + se)) +
    geom_pointrange() +
    facet_wrap(~ school_name)
```

## Regression (Table Three)
```{r regression-table}
#linear model of daily offenses counts on home games , away game indicators controlling for agency f
#ixed effects, day of week, month, and year. 
#reference
 daily_crime <- plot %>% group_by(date,home_or_away) %>% summarize (offense = sum (assault+vandalism) ) %>% filter(date == "2000-09-30" ) 
 
 regression_model <- plot %>% mutate(month =  month(date, label = TRUE, abbr = FALSE),
                                    year =   as.numeric(format(plot$date,"%Y") ) ,
                                    home = ifelse(home_or_away == "home",1,0),
                                    away = ifelse(home_or_away == "away",1,0)
 )
assault <- lm( assault ~ home+away+factor(ori) + factor(weekday)+as.character(month)+factor(year),data = regression_model)
summary(assault)
vandalism <- lm( vandalism ~ home+away+factor(ori) + factor(weekday)+factor(as.character(month))+factor(year),data = regression_model)
summary(vandalism)
coef(assault)["home"]
#outcome variable(dependent variable) = offenses , predictors = home_or_away 
#9% increase 
home_assault <- coef(assault)["home"] / mean(regression_model$assault)
home_assault
home_assault <- home_assault *100
home_assault
#not statistically sig
home_vandalism <-  coef(vandalism)["home"] / mean(regression_model$vandalism)
home_vandalism
home_vandalism <- home_vandalism * 100
home_vandalism
#not statistically sig
home_assault2 <- coef(assault)["away"] / mean(regression_model$assault)
home_assault2
home_assault2 <- home_assault2 *100
home_assault2
home_vandalism2 <-  coef(vandalism)["away"] / mean(regression_model$vandalism)
home_vandalism2
home_vandalism2 <- home_vandalism2 * 100
home_vandalism2
#away games show that it is there is no statsistical signifance and moreover no evidence to prove that there is a significance between offenses and away games 
# negative values so no effect 
regression_table1 <- broom::tidy(assault) %>% filter (term == "home" | term =="away") %>% mutate(percentage =  (coef(assault)[term] / mean(regression_model$assault)*100))%>% mutate (model  = "Assault")
regression_table2 <- broom::tidy(vandalism)%>% filter (term == "home" | term =="away") %>% mutate(percentage = (coef(vandalism)[term] / mean(regression_model$vandalism)*100)) %>% mutate(model = "vandalism")


regression_table <- bind_rows(regression_table1,regression_table2)
regression_table %>% pivot_wider(names_from = model, values_from = c(estimate, std.error, percentage ))
#possible cause of lower vandalism percenatge offense count maybe theirs is more , the paper uses negative =binomial coefficient while 
# this replication analysis use a linear regression moreover they use a larger data seta and the paper uses a larger set of agencies 
```
